---
title: "STAT547 Homework 1"
author: "Xingche Guo"
date: "9/8/2019"
output: pdf_document
---

# Problem 1 

```{r   echo = FALSE, warning = FALSE, message = FALSE }

library(fda)
library(ggplot2)
library(plotly)
library(reshape2)
library(RColorBrewer)
```

```{r}
data(pinch)
dim(pinch)

##### (b) #####
pinch_mu <- apply(pinch, 1, mean)
pinch_sd <- apply(pinch, 1, sd)

r_pinch <- melt(pinch)
names(r_pinch) <- c("time", "rep", "value")
r_pinch$time <- rep(pinchtime, 20)
r_pinch$rep <- as.factor(r_pinch$rep)

## raw
ggplot(data = r_pinch) + 
  geom_line(aes(x = time, y = value, group = rep), 
              alpha = 0.5) +
  geom_line(data = data.frame(x = pinchtime, y = pinch_mu), 
            aes(x=x, y=y), colour = "red", size = 1.5) + 
  geom_line(data = data.frame(x = pinchtime, y = pinch_sd), 
            aes(x=x, y=y), colour = "blue", size = 1) +
  theme_bw()

## smoothed
ggplot(data = r_pinch) + 
  geom_smooth(aes(x = time, y = value, group = rep), 
              alpha = 0.5, span = 0.1, se = FALSE, col = "gray") +
  geom_line(data = data.frame(x = pinchtime, y = pinch_mu), 
            aes(x=x, y=y), colour = "red", size = 1.5) + 
  geom_line(data = data.frame(x = pinchtime, y = pinch_sd), 
            aes(x=x, y=y), colour = "blue", size = 1) +
  theme_bw()


##### (c) #####
pinch_Cov <- cov(t(pinch))
persp(x=pinchtime, y=pinchtime, z=pinch_Cov)
contour(x=pinchtime, y=pinchtime, z=pinch_Cov)

#plot_ly(x=pinchtime, y=pinchtime, z=pinch_Cov) %>% add_surface(
#  contours = list(
#    z = list(
#      show=TRUE,
#      usecolormap=TRUE,
#      highlightcolor="#ff0000",
#      project=list(z=TRUE)
#    )
#  )
#)

```





# Problem 2

```{r  warning = FALSE, }
######################## Problem 2 #######################
D2 <- read.csv("/Users/apple/Desktop/ISU 2019 fall/STAT547/data/DataSets/Dow_companies_data.csv")

##### (a) #####
P1 <- D2$XOM[1]
date <- as.Date(as.character(D2$Date), format = "%m/%d/%y")

D_XOM <- data.frame(date = date, v1 = D2$XOM, 
                    v2 = 100*((D2$XOM/P1)-1))

ggplot(data = D_XOM) +
  geom_line(aes(x = date, y = v1)) + 
  ylab("stock value") + 
  ggtitle("Exxon-Mobil (XOM)") + theme_bw()

ggplot(data = D_XOM) +
  geom_line(aes(x = date, y = v2)) + 
  ylab("cumulative return function (%)") + 
  ggtitle("Exxon-Mobil (XOM)") + theme_bw()


D_XOM$v2[length(D_XOM$v2)]



##### (b) #####
D2_1 <- D2[,-1]

D2_2 <- apply(D2_1, 2, FUN = function(x){
  return( 100*(x - x[1])/x[1] )
})

r_D2_2 <- melt(D2_2)
names(r_D2_2) <- c("date", "stock", "CRF")
r_D2_2$date <- rep(date, 30)

stock_mean <- apply(D2_2, 1, mean)
stock_med <- apply(D2_2, 1, median)

DD <- data.frame(date = rep(date,2), 
                 CRF = c(stock_mean, stock_med),
                 label = rep(c("mean", "median"), each = 252))


ggplot(data = r_D2_2) + 
  geom_line(aes(x = date, y = CRF, group = stock), colour = "gray") + 
  geom_line(data = DD, aes(x = date, y = CRF, colour = label), size = 1) + 
  theme_bw()



##### (c) #####
fbplot(fit = D2_2, ylim = c(-20,90), xlab="Trading day",
       ylab = "Cumulative return")


 fbplot(fit = D2_2, ylim = c(-20,90), prob = c(0.9, 0.6, 0.5, 0.3),
       color = brewer.pal(4, "OrRd"),
       xlab="Trading day", 
       ylab = "Cumulative return")


```


# Problem 3

Suppose $\mathrm{Var}(\pmb{x}) = \Sigma = P \Lambda P^T$, where $\Lambda$ is the eigenvalue matrix such that $\lambda_1 \ge \lambda_2 \ge \dots \ge 0$, $P$ is the corresponding eigenvector matrix, then $\mathrm{Var}(\pmb{l}_1^T \pmb{x}) = \pmb{l}_1^T \Sigma \pmb{l}_1 = \pmb{y}_1^T \Lambda \pmb{y}_1$, where $\pmb{y}_1 = P^T \pmb{l}_1$, and $\pmb{y}_1^T \pmb{y}_1 = \pmb{l}_1^T \pmb{l}_1 = 1$. Note that:
$$
\pmb{y}_1^T \Lambda \pmb{y}_1 = \sum_{i=1}^d \lambda_i y_{1i}^2 \ge \lambda_1,
$$
equality holds iff $\pmb{y}_1 = \pmb{e}_1 = (1, 0, \dots, 0)^T$. Therefore, $\pmb{v}_1 = P\pmb{e}_1 = \pmb{p}_1$, where $\pmb{p}_1$ is the eigenvector corresponding to $\lambda_1$.

Also note that $\pmb{e}_1^T \pmb{y}_2 = \pmb{v}_1^T \pmb{l}_2 = 0$, which means $y_{21}=0$, 
therefore:
$$
\pmb{y}_2^T \Lambda \pmb{y}_2 = \sum_{i=2}^d \lambda_i y_{2i}^2 = \lambda_2,
$$

and $\pmb{v}_2 = P\pmb{e}_2 = \pmb{p}_2$, where $\pmb{p}_2$ is the eigenvector corresponding to $\lambda_2$. 

Thus, $V:=(\pmb{v}_1, \dots, \pmb{v}_d) = P$.


# Problem 4

## (a)
$$
\mathrm{E}(\xi_k) = \mathrm{E}(\pmb{v}_k^T (\pmb{x}-\pmb{\mu})) = 
\pmb{v}_k^T (\pmb{\mu}-\pmb{\mu}) = 0
$$

## (b)
$$
\mathrm{Var}(\xi_k) = \pmb{v}_k^T \Sigma  \pmb{v}_k = \pmb{e}_k^T P^T P \Lambda P^T P \pmb{e}_k = \lambda_k
$$

## (c)
$$
\mathrm{Cov}(\xi_j, \xi_k) = \pmb{v}_j^T \Sigma  \pmb{v}_k = \pmb{e}_j^T \Lambda  \pmb{e}_k = \delta_{jk} \lambda_k
$$

## (d)
$$
\mathrm{Corr}(X_j, \xi_k) = \frac{\mathrm{Cov}(X_j, \xi_k)}{\sigma_{jj} \sqrt{\lambda_k}} =  \frac{\pmb{e}_j^T \Sigma \pmb{v}_k}{\sigma_{jj} \sqrt{\lambda_k}} = \frac{\pmb{e}_j^T (\lambda_k \pmb{v}_k)}{\sigma_{jj}\sqrt{\lambda_k}} = \frac{v_{jk}\sqrt{\lambda_k}}{\sigma_{jj}}
$$


# Problem 5

## (a)
$$
||ax|| = <ax,ax>^{1/2} = |a| <x,x>^{1/2} = |a| \ ||x||
$$

## (b)
$$
\because \ <x + ay, x + ay> \ge 0, \ \ \ \ \forall a,
$$
$$
\therefore \ <y,y> a^2 + 2<x,y>a + <x,x> \ge 0, \ \ \ \ \forall a,
$$
$$
\therefore 4<x,y>^2 -4<x,x><y,y> \ge 0, 
$$
$$
\therefore |<x,y>| \le ||x|| \ ||y||.
$$

## (c)
$$
\because||x+y||^2 = <x+y, x+y> = <x,x> + <y,y> + 2<x,y> \le ||x||^2 + ||y||^2 + 2 ||x|| \ ||y||
=(||x|| + ||y||)^2,
$$
$$
\therefore ||x+y|| \le ||x|| + ||y||.
$$


## (d)

$$
d(x,y)=||x-y|| = <x-y, x-y>^{1/2} = <y-x, y-x>^{1/2} = ||y-x|| = d(y,x) > 0 , \ \ \ \ iff \ \ x\neq y,
$$
$$
d(x,x)=||x-x||=0
$$
$$
d(x,y)=||x-y|| = ||(x-z) + (z-y)|| \le ||x-z|| + ||z-y||=d(x,z)+d(z,y).
$$



# Problem 6

It's easy to find that $l_2$ space is a valid inner product space, we only need to verify that this space is complete. $\forall$ Cauchy sequence $\{x^{(n)}\}_{n=1}^{\infty}$, we have:
$$
lim_{n,m \rightarrow\infty}||x^{(n)} - x^{(m)}|| = 0,
$$
therefore, for large $n_k$, $\exists n_{k+1} > n_k$, such that:
$$
|| x^{(n_{k+1})} - x^{(n_k)} || \le \frac{1}{k^2}
$$
$$
\therefore \sum_{k=1}^{\infty} ||  x^{(n_{k+1})} - x^{(n_k)} || < \infty
$$
Define $x = \liminf_{k \rightarrow \infty} x^{(n_k)}$, we only need to prove (i). $||x||^2 < \infty$,
and (ii). $\lim_{n \rightarrow \infty} ||x - x^{(n)}|| = 0$.
First note that:
$$
||x||^2 = \sum_{i=1}^{\infty} |x_i|^2 = \sum_{i=1}^{\infty} \liminf_k |x_i^{(n_k)}|^2 \le \liminf_k \sum_{i=1}^{\infty} |x_i^{(n_k)}|^2 = \liminf_k || x^{(n_k)}||^2,
$$
The inequality comes from Fatou's lemma, also:
$$
\liminf_K ||x^{(n_K)}|| \le ||x^{(n_1)}|| + \liminf_K \sum_{k=1}^{K-1} || x^{(n_{k+1})} - x^{(n_k)} || < \infty,
$$
therefore $||x||^2 < \infty$.


Again, by Fatou's lemma:
$$
\lim_{n \rightarrow \infty} ||x - x^{(n)}||^2 = \lim_{n \rightarrow \infty} \sum_{i=1}^{\infty} |x_i - x_i^{(n)}|^2 = \lim_{n \rightarrow \infty} \sum_{i=1}^{\infty} \liminf_{k} |x_i^{(n_k)} - x_i^{(n)}|^2 \le \lim_{n ,k \rightarrow\infty} \sum_{i=1}^{\infty} |x_i^{(n_k)} - x_i^{(n)}|^2
$$
By the property of Cauchy sequence:
$$
\lim_{n \rightarrow \infty} ||x - x^{(n)}||^2 \le \lim_{n ,k \rightarrow\infty} ||x^{(n_k)} - x^{(n)}||^2 = 0,
$$
hence, $\lim_{n \rightarrow \infty} ||x - x^{(n)}|| = 0$. Thus, $l_2$ space to that end is an Hilbert Space.



# Problem 7


\begin{align*}
\mathrm{E}|| X - \sum_{k=1}^K <X, e_k> e_k ||^2 & = \mathrm{E} \int X^2(t) dt - 2\sum_{k=1}^{K} \mathrm{E}<X, e_k>^2 + \sum_{k=1}^K \sum_{j=1}^K \mathrm{E}<X, e_j> <X,e_k> \delta_{jk} \\
& = \int \mathrm{E} X^2(t) dt - \sum_{k=1}^{K} \mathrm{E}<X, e_k>^2 \\
& = \int G(t,t)dt - \sum_{k=1}^K \int \int e_k(s) e_k(t) G(s,t) ds dt.
\end{align*}

Thus, minimize $\mathrm{E}|| X - \sum_{k=1}^K <X, e_k> e_k ||^2$ is equivalent to maximize $\sum_{k=1}^K \int \int e_k(s) e_k(t) G(s,t) ds dt$. By the theorem, for K orthonormal basis $\{e_k\}_{k=1}^K$, we have:
$$
\max \sum_{k=1}^K \int \int e_k(s) e_k(t) G(s,t) ds dt = \sum_{k=1}^K \lambda_k,
$$
and
$$
\mathrm{argmax} \sum_{k=1}^K \int \int e_k(s) e_k(t) G(s,t) ds dt = \{\phi_k\}_{k=1}^{K},
$$
where $(\lambda_k, \phi_k)$ is the $k$th eigenvalue-eigenfunction pairs for NND function $G(s,t)$.



