---
title: "STAT547 Homework 2"
author: "Xingche Guo"
date: "9/23/2019"
output: pdf_document
---

# Problem 1

```{r}
library(fda)
library(fdapace)

############### Problem 1 #################
data(gait)
Hip <- gait[, , 1]
Knee <- gait[, , 2]
time <- as.numeric( rownames(Hip) )
n <- ncol(Hip)
m <- nrow(Hip)

XX <- t(rbind(Hip, Knee))
mu <- colMeans(XX) 
GG <- cov(XX)

## plot the mean function
par(mfrow = c(1,2))
matplot(time, Hip, type='l', col='black', main = "Hip Angle", xlab = "time") 
lines(time, mu[1:m], col = "red", lwd = 2.5)

matplot(time, Knee, type='l', col='black', main = "Knee Angle", xlab = "time") 
lines(time, mu[(m+1):(2*m)], col = "red", lwd = 2.5)
par(mfrow = c(1,1))

## check the total variation
eig <- eigen(GG)
lam <- eig$values / m 
FVEeach <- lam / sum(lam) 
FVE <- cumsum(FVEeach) 

plot(FVE, pch = 16)
abline(h = 0.95, col = "red")

## plot the eigen-functions
phi <- eig$vectors * sqrt(m)
phi1 <- phi[1:m,]
phi2 <- phi[(m+1):(2*m),]

par(mfrow = c(1,2))
matplot(time, phi1[,1:2], type='l', col=c('red', 'blue'), xlab = "time", ylab = "hip") 
matplot(time, phi2[,1:2], type='l', col=c('red', 'blue'), xlab = "time", ylab = "knee") 
par(mfrow = c(1,1))

## Mode of variation plot 
par(mfcol = c(2,2))
lamphi11 <- sqrt(lam[1]) * phi1[, 1] 
plot(time, mu[1:m], type='l', lwd=2, ylim = c(-10, 60), 
     main = "FPC1", xlab = "time", ylab = "hip") 
lines(time, mu[1:m] + lamphi11, type='l', lty=2, col='red') 
lines(time, mu[1:m] - lamphi11, type='l', lty=2, col='blue') 

lamphi12 <- sqrt(lam[2]) * phi1[, 2] 
plot(time, mu[1:m], type='l', lwd=2, ylim = c(-10, 60), 
     main = "FPC2", xlab = "time", ylab = "hip") 
lines(time, mu[1:m] + lamphi12, type='l', lty=2, col='red') 
lines(time, mu[1:m] - lamphi12, type='l', lty=2, col='blue') 

lamphi21 <- sqrt(lam[1]) * phi2[, 1] 
plot(time, mu[(m+1):(2*m)], type='l', lwd=2, ylim = c(-0, 80), 
     main = "FPC1", xlab = "time", ylab = "knee") 
lines(time, mu[(m+1):(2*m)] + lamphi21, type='l', lty=2, col='red') 
lines(time, mu[(m+1):(2*m)] - lamphi21, type='l', lty=2, col='blue') 

lamphi22 <- sqrt(lam[2]) * phi2[, 2] 
plot(time, mu[(m+1):(2*m)], type='l', lwd=2, ylim = c(-0, 80), 
     main = "FPC2", xlab = "time", ylab = "knee") 
lines(time, mu[(m+1):(2*m)] + lamphi22, type='l', lty=2, col='red') 
lines(time, mu[(m+1):(2*m)] - lamphi22, type='l', lty=2, col='blue') 
par(mfrow = c(1,1))


## FPCs
XXcenter <- XX - matrix( rep(mu, n), nrow=n, ncol=2*m, byrow=TRUE) 
xi <- XXcenter %*% phi / m 

plot(xi[, 1], xi[, 2], xlab = "FPC1", ylab = "FPC2", 
     col = "red", cex = 0.001) 
text(xi[, 1], xi[, 2], label= sub("boy", "", rownames(XX))) 

```


# Problem 2

```{r}

############### Problem 2 #################
K <- 20 
n <- 50 
m <- 101 
t <- seq(0, 1, length.out = m) 
mu <- -(t - 0.5)^2 + 1 


## Create Basis 
phiSim <- CreateBasis(K=K, t, type='fourier') 
#matplot(t, phiSim[, 1:5], type='l') 
lamSim <- exp(-(1:K)) 

## Gaussian Process
# rows are for individuals, col are for different FPC 
xi <- matrix(rnorm(n * K), nrow = n, ncol = K) %*% diag(sqrt(lamSim)) 
X1 <- matrix(mu, nrow=n, ncol=m, byrow=TRUE) + xi %*% t(phiSim) 
matplot(t, t(X1), type='l') 


## non-Gaussian Process
xi <- matrix(rt(n * K, df = 2), nrow = n, ncol = K) %*% diag(sqrt(lamSim)) 
X2 <- matrix(mu, nrow=n, ncol=m, byrow=TRUE) + xi %*% t(phiSim) 
matplot(t, t(X2), type='l') 

```


# Problem 3

```{r}

yeast <- read.table(file = "/Users/apple/Desktop/ISU 2019 fall/STAT547/data/yeast.txt")
time <- as.numeric( sub("alpha", "", names(yeast)) )

## original data
X <- as.matrix(yeast)
matplot(time, t(X), type='l', col=c(rep(2,44), rep(4,45)), 
        xlab = "time", ylab = "yeast") 

n <- nrow(X)
m <- ncol(X)

## smoothed data
X_smooth <- matrix(0, ncol = m, nrow = n)

for (i in 1:n){
  ind <- which(is.na(X[i,]))
  if (length(ind) > 0){
    fit <- smooth.spline(time[-ind], X[i,][-ind])
  }else{
    fit <- smooth.spline(time, X[i,], all.knots = TRUE, spar = 0.1)
  }
    xhat <- predict(fit, as.data.frame(time))$y$time
    X_smooth[i,] <- xhat
}

matplot(time, t(X_smooth), type='l', col=c(rep(2,44), rep(4,45)), 
        xlab = "time", ylab = "yeast") 


## check the total variation
mu <- colMeans(X_smooth) 
G <- cov(X_smooth) 
eig <- eigen(G) 
lam <- eig$values * diff(range(time)) / m 
FVEeach <- lam / sum(lam) 
FVE <- cumsum(FVEeach) 

plot(FVE, pch = 16)
abline(h = 0.95, col = "red")


## Mode of variation plot 
phi <- eig$vectors / sqrt(diff(range(time) / m)) 

par(mfrow = c(2,2))
lamphi1 <- sqrt(lam[1]) * phi[, 1] 
plot(time, mu[1:m], type='l', lwd=2, ylim = c(-1,1),
     main = "FPC1", xlab = "time", ylab = "PC1") 
lines(time, mu + lamphi1, type='l', lty=2, col='red') 
lines(time, mu - lamphi1, type='l', lty=2, col='blue') 

lamphi2 <- sqrt(lam[2]) * phi[, 2] 
plot(time, mu, type='l', lwd=2, ylim = c(-1, 1), 
     main = "FPC2", xlab = "time", ylab = "PC2") 
lines(time, mu + lamphi2, type='l', lty=2, col='red') 
lines(time, mu - lamphi2, type='l', lty=2, col='blue') 

lamphi3 <- sqrt(lam[3]) * phi[, 3] 
plot(time, mu, type='l', lwd=2, ylim = c(-1, 1), 
     main = "FPC3", xlab = "time", ylab = "PC3") 
lines(time, mu + lamphi3, type='l', lty=2, col='red') 
lines(time, mu - lamphi3, type='l', lty=2, col='blue') 

lamphi4 <- sqrt(lam[4]) * phi[, 4] 
plot(time, mu, type='l', lwd=2, ylim = c(-1, 1), 
     main = "FPC4", xlab = "time", ylab = "PC4") 
lines(time, mu + lamphi4, type='l', lty=2, col='red') 
lines(time, mu - lamphi4, type='l', lty=2, col='blue') 
par(mfrow = c(1,1))


## FPC
Xcenter <- X - matrix(mu, nrow=n, ncol=m, byrow=TRUE) 
xi <- Xcenter %*% phi * diff(range(time)) / m 
plot(xi[, 1], xi[, 2],  col=c(rep(2,44), rep(4,45)), pch = 16,
     xlab = "FPC1", ylab = "FPC2") 

```


# Problem 4

## (a)

$$
Z_k = \frac{1}{n} \sum_{i=1}^n  \int_{\mathcal{T}} (X_i(t) - \hat{\mu}(t)) \hat{\phi_k}(t) dt = 
  \int_{\mathcal{T}} ( \frac{1}{n} \sum_{i=1}^nX_i(t) - \hat{\mu}(t)) \hat{\phi_k}(t) dt = 0
$$

## (b)


\begin{align*}
& \frac{1}{n-1} \sum_{i=1}^n (\hat{\xi}_{ik} - Z_k)^2 = \frac{1}{n-1} \sum_{i=1}^n \hat{\xi}_{ik}^2
= \frac{1}{n-1} \sum_{i=1}^n \int_{\mathcal{T}}\int_{\mathcal{T}} \hat{\phi}_k(s) (X_i(s) - \hat{\mu}(s)) (X_i(t) - \hat{\mu}(t)) \hat{\phi}_k(t) ds dt \\
= & \int_{\mathcal{T}}\int_{\mathcal{T}} \hat{\phi}_k(s) \hat{G}(s,t) \hat{\phi}_k(t) ds dt =
\int_{\mathcal{T}} \hat{\phi}_k(s) \hat{\phi}_k(s) \hat{\lambda}_k ds = \hat{\lambda}_k
\end{align*}



## (c)

\begin{align*}
& \frac{1}{n-1} \sum_{i=1}^n (\hat{\xi}_{ik} - Z_k)(\hat{\xi}_{ik^{'}} - Z_{k^{'}}) = \frac{1}{n-1} \sum_{i=1}^n \hat{\xi}_{ik} \hat{\xi}_{ik^{'}} =
\int_{\mathcal{T}}\int_{\mathcal{T}} \hat{\phi}_k(s) \hat{G}(s,t) \hat{\phi}_k(t) ds dt \\
= & \int_{\mathcal{T}} \hat{\phi}_{k^{'}}(s) \hat{\phi}_k(s) \hat{\lambda}_{k^{'}} ds =
 \hat{\lambda}_k \delta_{k k^{'}}
\end{align*}




# Problem 5

Define: 
$\mathcal{J} = \sum_{i=1}^n (x_i - \bar{x}) \otimes  (x_i - \bar{x})$, then $\forall y \in \mathbb{H}$:

$$
\mathcal{J}(y) = \sum_{i=1}^n (x_i - \bar{x}) <x_i - \bar{x}, y>,
$$
which means $\mathcal{Im}(\mathcal{J}) = \mathrm{Span\{ x_i - \bar{x} \}}_{i=1}^n$. Also note that $\{x_i - \bar{x}\}_{i=1}^n$ is one degree of freedom less than $n$ sine $\sum_{i=1}^n x_i - \bar{x} = 0$, therefore:

$$
\mathrm{Rank}(\mathcal{J}) = \mathrm{dim}(\mathcal{Im}(\mathcal{J})) = n-1.
$$


# Problem 6

## (a)

To show absolute convergence, we only need to prove:
$$
\sum_{j=1}^{\infty} || a_j \lambda_j e_j ||_K < \infty,
$$
This is true because:
$$
\sum_{j=1}^{\infty} || a_j \lambda_j e_j ||_K = \sum_{j=1}^{\infty}  <a_j \lambda_j e_j, a_j \lambda_j e_j>_K = \sum_{j=1}^{\infty} a_j^2 \lambda_j < \infty.
$$


## (b)
Because:
$$K(\cdot, t) = \sum_{j=1}^{\infty} \lambda_j e_j(\cdot) e_j(t),$$
Let $a_{t,j} = e_j(t)$, then:
$$
\sum_{j=1}^{\infty} \lambda_j a_{t,j}^2 = \sum_{j=1}^{\infty} \lambda_j e_j(t) e_j(t) = 
K(t, t) < \infty,
$$
The last inequality holds because $\{K(t,t), t \in [0,1]\}$ is compact. 


## (c)
$\forall f \in \mathcal{H}_K$, $\exists f = \sum_{j=1}^{\infty} a_j \lambda_j e_j$, therefore:
$$
<K(\cdot, t), f>_K = <\sum_{j=1}^{\infty} \lambda_j e_j(t) e_j, \sum_{j=1}^{\infty} a_j \lambda_j e_j>_K = \sum_{j=1}^{\infty} \lambda_j a_j e_j(t) = f(t).
$$















